---
title: "tester package"
author: "Craig Parman"
date: "3/16/2015"
output: pdf_document
--- 


###Introduction

The tester package provise a set of east to read tests that return logical vector.  The test can be used in conjunction with other testing packages to develop very powerful, yet easy to read tests.  For examnple to test if a function return a floting point number less than zero one woulf need an expression like
```{r} 
    number = -0.56

    (number < 0 && is.numeric(number) )

```

With tester the test is much more readable and friendly code.

```{r}
library(tester)

number = -0.56

is_negative_decimal(number) 

```

The package also allows you to easily write tests that look for a single condition across a group of items.  For example if I hve a vector of numerics and I want check is any are infinite values.  rather than teste each element individually and checking to see if any test have failed, I can do one test and get a single logical result.
```{r}
library(tester)

n<-c(rnorm(10),1/0)

n

has_infinite(n)

```

You can also test for attributes of an object such as dimensions, dimnames, and 
missing values. For example if you needed to test whether an object was a numeric matrix with dimension names you could construct the following test.

```{r}

m<-matrix(1:16,nrow=4,ncol=4)
dimnames(m)<-list(rep("row",4),rep("col",4))

is_numeric_matrix(m) & has_dimnames(m)

```




### Catalog of Tests

As of version 0.1.7 the tester package has over 100 tests. The tests all have the form 'type of comparision'\_'comparision' such as in the example above, is_numeric().  

There are 5 different comparision types and many diffrent comparisions and combinations of comparisions.

Comparision Types 

is                
has               
different         
lacks             
same              


These are the 'is' tyep test with numeric comparisions.
 

test                          |  Description
----------------------------- | -----------------------------------------------
is_even                       | Is even
is_not_even                   | Is not even
is_odd                        | Is odd
is_not_odd                    | Is not odd
is_positive                   | Is positive
is_not_positive               | Is not positive
is_negative                   | Is negative
is_not_negative               | Is not negative
is_decimal                    | Is decimal
is_not_decimal                | Is not decimal
is_integer                    | Is integer
is_not_integer                | Is not integer
is_natural                    | Is natural
is_not_natural                | Is not natural
is_positive_decimal           | Is positive decimal
is_negative_decimal           | Is negative decimal
is_positive_integer           | Is positive integer
is_negative_integer           | Is negative integer
is_positive_scalar            | Is postive scalar
is_negative_scalar            | Is negatice scalar
is_multiple                   | Is x a multiple of y


These tests are for object class and dimensions.



test                          |  Description
----------------------------- | -----------------------------------------------
is_class                      | x is of class y
is_dataframe                  | Is a data.frame
is_not_dataframe              | Is not  data.frame
is_factor_dataframe           | Is a data.frame of factors
is_string                     | Is string
is_not_string                 | Is not string
is_string_tabular             | Is string tabular
is_string_vector              | Is string vector
is_string_dataframe           | Is string data.frame
is_tabular                    | Is tabular
is_not_tabular                | Is not tabular
is_numeric_dataframe          | Is numeric data.frame
is_numeric_tabular            | Is numeric tabular
is_numeric_vector              | Is numeric vector
is_vector                     | Is vector
is_not_vector                 | Is not vector
is_scalar                     | Is scalar
is_not_scalar                 | Is not scalar
is_one_dim                    | Test if an object has one-dimension


These are all of the tests related to matrix objects.

test                         |  Description
---------------------------- | ------------------------------------------------
is_matrix                    | Is matrix
is_not_matrix                | Is not matrix
is_numeric_matrix            | Is numeric matrix
is_string_matrix             | Is string matrix
is_square_matrix             | Is square matrix
is_not_square_matrix         | Is not square matrix
is_square_numeric_matrix     | Is square numeric matrix
is_not_square_numeric_matrix | Is not square numeric matrix
is_diagonal                  | Is diagonal matrix
is_not_diagonal              | Is not diagonal matrix
is_triangular_matrix         | Is triangular matrix
is_lower_triangular          | Is triangular matrix
is_upper_triangular          | Is triangular matrix




These are the logical tests.

test                         |  Description
---------------------------- | ------------------------------------------------
is_TRUE                      | Is TRUE
is_true                      | Is TRUE
is_FALSE                     | Is FALSE
is_false                     | Is FALSE
is_logical_matrix            | Is matrix of logicals
is_logical_vector            | Is vector of logicals
true_or_false                | whether TRUE or FALSE, False for non logicals





These is a series of tests rthat test the the object has a single value ao different types and atrributes.

test                         |  Description
---------------------------- | ------------------------------------------------
is_single                    | Is single
is_single_decimal            | Is single decimal
is_single_even               | Is single even
is_single_false              | Is single false
is_single_logical            | Is single logical
is_single_negative           | Is single negative number
is_single_negative_decimal   | Is single negative decimal
is_single_negative_integer   | Is single negative integer
is_single_number             | Is single number
is_single_odd                | Is single odd
is_single_positive           | Is single positive number
is_single_positive_decimal   | Is single positive decimal
is_single_positive_integer   | Is single positive integer
is_single_string             | Is single string
is_single_true               | Is single true



The 'has' test test for various object atrtributes and missing or non-numeric 
values.

test                          |  Description
----------------------------- | -----------------------------------------------
has_colnames                  | Has column names
has_rownames                  | Has row names
has_dim                       | Has dimension
has_dimension                 | Has dimension
has_dimnames                  | Has dimesion names
has_factors                   | Has factors
has_Inf                       | Has Inf or -Inf values
has_infinite                  | Has Inf or -Inf values
has_missing                   | Has NA values 
has_NA                        | Has NA values 
has_names                     | Has names
has_NaN                       | Has  NaN values
has_not_a_number              | Has NaN values
has_nas                       | Has any NA, NaN, Inf values


The following test compare objects.  These are of the form test(x,y)

```{r}
different_length(rep(1,4),rep(1,3))  
```

test                          |  Description
----------------------------- | -----------------------------------------------
same_class                    | Same Class
different_class               | Different Class
same_dim                      | Same Dimension
different_dim                 | Different Dimension
same_length                   | Same Length
different_length              | Different Length
same_mode                     | Same Mode
different_mode                | Different Mode
same_ncol                     | Same Number of Columns
different_ncol                | Different Number of Columns
same_nrow                     | Same Number of Rows
different_nrow                | Different Number of Rows
same_type                     | Same Type
different_type                | Different Type











