---
title: "quickcheck Package"
author: "Craig Parman"
date: "2/24/2015"
output: html_document
---

## *quickcheck* Package 


###Introduction

The package quickcheckform [RevolutionAnalytics] http://www.revolutionanalytics.com adds a twist to unti testing. Rather than specifing exact values to be tested, set of random values of a particular type is passed.

Here is a simple example.  Let test thet the funtion returns a numeric value.

A simple test using testthat might look like:

```{r}
library(testthat)

test_that("rest for floor", expect_identical(is.numeric(floor(3.14)), TRUE))


```

Here we specify the exact value 3.14 to be passed to the function floor.  With quickcheck you don't specify the value, but rather the class of a set of random variables to be passed to the function.

```{r}
library(quickcheck)

test(function(x = rdouble())  identical(is.numeric(floor(x)), TRUE))


```

The test is not being run on a single value for x, but a series of values generated by rdouble.  In fact you not only get a random set of doubles, you get a randon number of values.  To see this lets run rdouble twice.  We will set the random number seed to 0 for reproducibility.

```{r}
library(quickcheck)

set.seed(0)
rdouble()

rdouble()

```

 
In the first case we get 13 values and in the second we get 9.  We can further control the number generation with the element and size parameters. The element value sets expectation of the values and size sets the average number of values returned as displayed below.


```{r}
library(quickcheck)

set.seed(0)

set1 <- rdouble(element=100,size=10)

set2<-rdouble(element=10,size=100)

mean(set1)
length(set1)

mean(set2)
length(set2)

```



The power of quickcheck comes from you can pass a



####Usage

###Examples
